---
title: 'Python vs R for classification '
author: Harry Fisher
date: '2019-08-19'
slug: Python-vs-r-for-machine-learning
draft: true 
categories:
  - Python
  - R
  - Machine Learning
tags:
  - Python
  - R
subtitle: ''
summary: 'Exploring machine learning techniques in different languages :robot:'
authors: []
lastmod: '2019-08-16T18:34:59+01:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
output: distill::distill_article
---

I often see comments about Python vs R for "insert any kind of task here". They are often treated like arch enemies in an internal struggle for superiority where there can only be one winner. 
While they are of course very different languages, my take is that they are both extremely useful tools and have their own strengths and weaknesses and can be used interchangeably. 
Recently, I have been trying to expand my Python knowledge as I think it is important to have a range of tools available for any data science problem. 
While I have found R to be more than adequate for problems I have faced, I thought it would be good practice to implement some different models using both languages.

In this post I will compare how R and Python can be used to address a simple classification problem using a couple of common techniques. 
Given my slight R bias, I am going to be using the [`recticulate`](https://cran.r-project.org/web/packages/reticulate/index.html) package to run Python code within R. 

I have chosen three different methods; logistic regression, random forests and Support Vector Machines (SVM) as these are widely used for many types of classification problems.

- In R, I'll be using the [`tidymodels`](https://www.tidyverse.org/articles/2018/08/tidymodels-0-0-1/) packages, including [`recipies`](https://cran.r-project.org/web/packages/recipes/index.html) and [`parsnip`](https://cran.r-project.org/web/packages/parsnip/index.html) for preparing the data and fitting the models and [`yardstick`](https://cran.r-project.org/web/packages/yardstick/index.html) for model evaluation.

- In Python, I'll be using the [Scikit-learn](https://scikit-learn.org/stable/tutorial/basic/tutorial.html) library which as many different modelling possibilities.


# The data

I'll be using is the credit dataset that comes with the recipes package. The data contains 4454 rows with several predictor variables including job, marital status and age, and the variable we'll be trying to predict, credit status, which can be 'good' or 'bad'. 

```{r loadthemin, message=FALSE}
library(tidyverse) 
library(tidymodels)
library(reticulate) # For using Python in R
library(janitor) # data cleaning
library(DataExplorer) # EZ data Viz
library(parsnip) # interface for different modelling methods
library(knitr)

use_condaenv("tensorflow") # pointing R to my conda environment "tensorflow" which has many appropriate libraries intalled
  
```

```{r data}
data("credit_data")
df <- credit_data %>% 
  clean_names()

head(df) %>% 
  kable()
```


I Like to use the [`DataExplorer`](https://cran.r-project.org/web/packages/DataExplorer/vignettes/dataexplorer-intro.html) package to perform some quick checks on the data.

```{r prep}
introduce(df) %>% 
  gather(key, value) %>% 
  kable()
```

This tells us there are some missing values, and a mix of continuous and categorical predictors meaning we will have to do some preprocessing.

Thankfully, the recipes package makes this very straightforward:

- First, we'll split the data into training and testing sets.
- Then we'll use the "step_X" functions in recipes to perform the necessary preprocessing.
- Before finally "baking" our processed data set that we can use for our models (a definite cooking them going on here...).


```{r prepro}
set.seed(123)
train_test_split <- initial_split(df)

df_train <- training(train_test_split)
df_test <- testing(train_test_split)

df_pro <- recipe(status ~ ., data = df) %>% 
  step_knnimpute(all_predictors()) %>% # missing values
  step_dummy(all_predictors(), -all_numeric()) %>% # dummify categoricals
  step_center(all_predictors()) %>% # center 
  step_scale(all_predictors()) # scale

trained_rec <- prep(df_pro, training = df_train)

train_data <- bake(trained_rec, new_data = df_train)
test_data  <- bake(trained_rec, new_data = df_test) 

```

# Logistic regression

First up is a simple logistic regression. 

- First we'll specify the model using the parsnip interface.
- Then we'll make predictions  using the model on our test data set
- We can then assess how well the model did using some common model assessment criteria

In R we can do...

```{r logreg}
set.seed(124)

model_log <- 
  logistic_reg() %>% 
  set_engine("glm") %>% 
  fit(status ~ ., data = train_data)

pred_log <- model_log %>% 
  predict(new_data = test_data) %>% 
  bind_cols(test_data %>% select(status))

# function to evaluate model performance
eval <- function(model){
  model %>% 
  conf_mat(.pred_class, status) %>%
  summary() %>%
  select(-.estimator) %>%
  filter(.metric %in%
    c("accuracy", "precision", "recall", "f_meas"))
  kable()
}

eval(pred_log)

```

The logistic regression model gave us 78 % accuracy, which is not bad considering how little effort we put into the modelling step.

For Python...

```{python logregpy, message = FALSE}
import numpy as np
import matplotlib.pyplot as plt
import pandas as pd

from sklearn.linear_model import LogisticRegression
from sklearn import metrics

# Get r data into Python

X = r.train_data  
y = r.test_data 

# Prepare training and testing sets

X_train = X.iloc[:, 1:23]
y_train = X.iloc[:,0]
X_test = y.iloc[:, 1:23]
y_test = y.iloc[:,0]

# Fit the model

logreg = LogisticRegression(solver='lbfgs', random_state = 124)
fit = logreg.fit(X_train, y_train)

# Predict

y_pred_log = logreg.predict(X_test)


from sklearn.metrics import classification_report

print(classification_report(y_test, y_pred_log))


```

Unsurprisingly the model gives very similar results with again an accuracy of 78%.

# Random Forest

Random forests are a popular ML technique for classification problems, and improve upon a single decision tree by combining many different trees to improve our predictions (for a great overview see [here](https://medium.com/@williamkoehrsen/random-forest-simple-explanation-377895a60d2d)).

In R, all we need to do is change the model specification and we're good to go. We can actually combine some of the code to reduce the number of lines it takes.
```{r rf}

set.seed(125)  

model_rf <- 
  rand_forest(mode = 'classification', trees = 1000) %>% 
  set_engine("ranger") %>% 
  fit(status ~ ., data = train_data) %>% 
  predict(new_data = test_data) %>% 
  bind_cols(test_data %>% select(status))

eval(model_rf)
```

Interesting, this model performs marginally worse than our logistic regression, but still rounds to 78% accuracy. There are many other techniques we could use to improve the model (e.g., cross validation) but for simplicity sakes we'll leave it as that for now.

In Python, we only need three lines of code:

```{python rfpy, message = FALSE}

from sklearn.ensemble import RandomForestClassifier

rf = RandomForestClassifier(n_estimators = 1000, random_state = 125)
f_fit = rf.fit(X_train, y_train)
y_pred_rf = rf.predict(X_test)

print(classification_report(y_test, y_pred_rf))

```

Again, very similar performance to the R model.


As a quick bonus we can take a look at the most important feastures in the data set.

```{python warning = FALSE}
feature_imp = pd.Series(rf.feature_importances_,index=list(X_train.columns)).sort_values(ascending=False)
print(feature_imp.head())
```

The output shows the top 5 most important features that are used to make predictions in the model. 

# Support Vector Machines

Support vector machines are another technique that can be used for classification. 

Defined [here](https://towardsdatascience.com/support-vector-machine-introduction-to-machine-learning-algorithms-934a444fca47):

> The objective of the support vector machine algorithm is to find a hyperplane in an N-dimensional space(N — the number of features) that distinctly classifies the data points.

Again, these can be easily implemented in both R and Python.

In R:

```{r svm, message=FALSE, warning=FALSE}
set.seed(127)

model_svm <- 
  svm_poly(mode = 'classification') %>% 
  set_engine("kernlab") %>% 
  fit(status ~., data = train_data) %>% 
  predict(new_data = test_data) %>% 
  bind_cols(test_data %>% select(status))

eval(model_svm)

```

Again in Python its a simple three-liner:

```{python svmpy}

from sklearn.svm import SVC

clf = SVC(kernel='linear', random_state = 127)
svm_fit = clf.fit(X_train, y_train)
y_pred_svm = clf.predict(X_test)

print(classification_report(y_test, y_pred_svm))

```

SVM performs slightly worse than both the logstic regression and random forest models, achieving an accuracy of 77%. However, like random forests, SVMs have several optimisation factors we could modify to improve the models performance.

# Summary

The point here was to just explore how different models could be specified in both R and Python using popular packages and libraries.
There are many other steps we could take to improve each models performance, but this post would be getting quite long if we optimised each example. 
I may do some plotting examples in R and Python in a future blog post, but until then, thanks for reading!



<!-- ```{Python echo = FALSE} -->

<!-- print('Logistic regression accuracy: {:.2f}'.format(logreg.score(X_test, y_test))) -->
<!-- print("Random Forest Accuracy: {:.2f}".format(metrics.accuracy_score(y_test, y_pred_rf))) -->
<!-- print("SVM Accuracy: {:.2f}".format(metrics.accuracy_score(y_test, y_pred_svm))) -->


<!-- ``` -->




